---
title: "Business Intelligence Lab Submission Markdown"
author: "Kelly Noella Sota"
date: "22/09/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                                   |                                                                                            |
|--------------------------|----------------------------------------------|
| **Student ID Numbers and Names of Group Members** | \<list one student name and ID per line; you should be between 2 and 5 members per group\> |

```         
                                                  Kenneth Kimathi
                                                  Kelly Noella Sota
```

| **BBIT 4.2 Group** \| \<either group A, B, or C\> \|
| **Course Code** \| BBT4206 \|
| **Course Name** \| Business Intelligence II \|
| **Program** \| Bachelor of Business Information Technology \|
| **Semester Duration** \| 21^st^ August 2023 to 28^th^ November 2023 \|

# Setup Chunk

We start by installing the required packages:

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
# Install renv:
if (!is.element("renv", installed.packages()[, 1])) {
  install.packages("renv", dependencies = TRUE)
}
require("renv")

```

------------------------------------------------------------------------

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

# Loading the dataset, check its data dimensions and data types

The dataset will be loaded into the project environment. The following code will show us the dataset dimension (number of rows/observations and columns/variables). And the last chuck of code will show the different data types of the variables.

```{r load dataset}
 #Load dataset
  
StudentPerformance_Dataset_1 <- read.csv("data/StudentPerformance_Dataset_1.csv", header = FALSE,
                         stringsAsFactors = TRUE)                                                                                                     
#Dimensions
dim(StudentPerformance_Dataset_1)

#Data types
sapply(StudentPerformance_Dataset_1, class)

```

# Dataset reduction

The following code will, firstly, reduce the dataset to a subset, check for its dimension and variables data types.

```{r dataset reduction}

subset_StudentPerformance <- StudentPerformance_Dataset_1 [,1:10]
dim(subset_StudentPerformance)
sapply(subset_StudentPerformance,class)

```

# Frequency and Central Tendency Measure

The frequency provides insights about the distribution of values within a specific variable, in our case the YOB (year of birth). While the central tendency provides the most frequent value of a variable (mode).

```{r Frequency, Central Tendency}
#Frequency Measure
SubsetStudentPerformance_freq <- subset_StudentPerformance$V3
cbind(frequency = table(SubsetStudentPerformance_freq),
      percentage = prop.table(table(SubsetStudentPerformance_freq)) * 100)

#Central Tendency Measure
SubsetStudentPerformance_V3_mode <- names(table(
    subset_StudentPerformance$V3))[
    which(table(subset_StudentPerformance$V3) ==
    max(table(subset_StudentPerformance$V3)))
]
print(SubsetStudentPerformance_V3_mode)
```

# Variability Measures

Standard deviation and variance can be calculated for numeric variables. The standard deviation shows how spread out the data values are, while the variance indicates how much each data point varies from the mean of the dataset.

```{r Standard deviation, Variance}
#Standard Deviation Measure
sapply(subset_StudentPerformance[, 7:10], sd)

#Variance Measure
sapply(subset_StudentPerformance[, 7:10], var)
```

# Kurtosis and Skewness

Kurtosis measures the degree to which the data distribution is peaked or flat in comparison to a normal distribution, while the skewness measures the lack of symmetry in the distributionof data.

```{r Kurtosis, Skweness}
#Kurtosis Measure
sapply(subset_StudentPerformance[, 7:10], kurtosis, type = 2)

#Skewness Measure
sapply(subset_StudentPerformance[, 7:10], skewness, type = 2)
```

# Optional Plots

Histograms and box plots have been created to represent the data distribution. Histogram represents the numeric variables while the the box plots will represent a summary of the distribution of numeric variables.

```{r Histograms, Box Plot}
#Histogram Creation

subset_StudentPerformance_YOB <- as.numeric(unlist(subset_StudentPerformance[, 3]))
hist(subset_StudentPerformance_YOB, main = names(subset_StudentPerformance)[3])

subset_StudentPerformance_read <- as.numeric(unlist(subset_StudentPerformance[, 7]))
hist(subset_StudentPerformance_read, main = names(subset_StudentPerformance)[7])

subset_StudentPerformance_anticipate <- as.numeric(unlist(subset_StudentPerformance[, 8]))
hist(subset_StudentPerformance_anticipate, main = names(subset_StudentPerformance)[8])

subset_StudentPerformance_answer <- as.numeric(unlist(subset_StudentPerformance[, 9]))
hist(subset_StudentPerformance_answer, main = names(subset_StudentPerformance)[9])

subset_StudentPerformance_find <- as.numeric(unlist(subset_StudentPerformance[, 10]))
hist(subset_StudentPerformance_find, main = names(subset_StudentPerformance)[10])



#Box and Whisker Plots Creation

subset_StudentPerformance_YOB <- as.numeric(unlist(subset_StudentPerformance[, 3]))
boxplot(subset_StudentPerformance_YOB, main = names(subset_StudentPerformance)[3])

subset_StudentPerformance_read <- as.numeric(unlist(subset_StudentPerformance[, 7]))
boxplot(subset_StudentPerformance_read, main = names(subset_StudentPerformance)[7])

subset_StudentPerformance_anticipate <- as.numeric(unlist(subset_StudentPerformance[, 8]))
boxplot(subset_StudentPerformance_anticipate, main = names(subset_StudentPerformance)[8])

subset_StudentPerformance_answer <- as.numeric(unlist(subset_StudentPerformance[, 9]))
boxplot(subset_StudentPerformance_answer, main = names(subset_StudentPerformance)[9])

subset_StudentPerformance_find <- as.numeric(unlist(subset_StudentPerformance[, 10]))
boxplot(subset_StudentPerformance_find, main = names(subset_StudentPerformance)[10])
```

**etc.** as per the lab submission requirements.
